### 2023-12-14 doc/manuscript.tex
update doc, directional deserts

\documentclass{article}
\usepackage[round]{natbib}

\usepackage[letterpaper,margin=1in]{geometry}
\usepackage{lineno}   %% <- no mathlines option
\usepackage{amsmath}  %% <- after lineno
\usepackage{etoolbox} %% <- for \cspreto, \csappto
\usepackage{amsfonts}
\usepackage{graphicx}
\usepackage{color}
\usepackage{url}

\usepackage{setspace, caption}
\captionsetup{font=onehalfspacing} % Spacing in float captions

\usepackage{xr}
\externaldocument{supplement}

\DeclareMathOperator{\Var}{Var}
\newcommand{\aprcomment}[1]{{\textcolor{blue}{APR: #1}}}
\newcommand{\E}{\mathbb{E}}
\usepackage{xspace}
\newcommand{\moments}{\texttt{moments}\xspace}
\newcommand{\fwdpy}{\texttt{fwdpy11}\xspace}
\newcommand{\tskit}{\texttt{tskit}\xspace}

%% Patch 'normal' math environments:
\newcommand*\linenomathpatch[1]{%
  \cspreto{#1}{\linenomath}%
  \cspreto{#1*}{\linenomath}%
  \csappto{end#1}{\endlinenomath}%
  \csappto{end#1*}{\endlinenomath}%
}

\linenomathpatch{equation}
\linenomathpatch{gather}
\linenomathpatch{multline}
\linenomathpatch{align}
\linenomathpatch{alignat}
\linenomathpatch{flalign}

%\linenumbers

\title{Archaic introgression and the distribution
of shared functional variation under stabilizing selection}
\author{APR}
\date{\today}

\begin{document}
\maketitle

\begin{abstract}

    Many phenotypic traits are under stabilizing selection, which maintains
    individual phenotypic values near some optimum. Classical theory shows that
    at loci contributing to the genetic variance of a trait under stabilizing
    selection, the minor allele is selected against via symmetric
    underdominance. Most studies of the dynamics of the genetic architecture
    and variance of a trait under stabilizing selection have focused on single
    populations. However, natural populations are typically structured in some
    way, and admixture and introgression are common occurances, in particular
    in human evolution. Because stabilizing selection results in selection
    against the minor allele at a trait-affecting locus, the alleles from the
    minor parental ancestry will be selected against after admixture. Here, we
    show that the site frequency spectrum can be used to model the genetic
    architecture of such traits, allowing for complex multi-population
    histories including admixture. We develop a simple model with linkage to a
    neutral locus to show that introgressed ancestry is expected to be depleted
    around trait-affecting alleles. Using this model and individual-based
    simulations, we show that introgressed ancestry deserts tend to form around
    such loci. When introgression between two diverged populations occurs in
    both directions, as has been inferred between humans and Neanderthals,
    introgressed ancestry deserts will often be shared across populations.
    Stabilizing selection for a shared phenotypic optimum may therefore explain
    recent observations in which regions of depleted human-introgressed
    ancestry in the Neanderthal genome overlap with Neanderthal-ancestry
    deserts in humans.

\end{abstract}

\onehalfspacing

\section*{Introduction}

Genomic surveys of natural systems show that historical admixture among
diverged populations or closely related taxa commonly occurs
\citep{brandvain2014speciation, skoglund2015ancient, suvorov2022widespread} and
is potentially widespread in primate \citep{tung2017contribution,
sorensen2023genome} and hominin \citep{wolf2018outstanding, peter2020100}
evolution. Admixture may therefore be an important driver of phenotypic and
molecular variation and shapes the genetic basis of complex traits. In humans,
archaic introgression involving Neanderthals and Denisovans has attracted
considerable attention, including efforts to describe the historical processes
leading to observed distributions of introgressed DNA in present-day
populations \citep{prufer2014complete, villanea2019multiple,
chen2020identifying} and the contribution of introgressed variation to
quantitative traits \citep{sankararaman2016combined, wei2023lingering}.

Once introduced through admixture, introgressed alleles may be selected for or
against. Some introgressed haplotypes were likely positively selected in modern
humans \citep{huerta2014altitude, racimo2017signatures, enard2018evidence,
gower2021detecting}, possibly due to locally adaptive variation that provided
fitness advantages as humans encountered novel environments. Despite some cases
of adaptive introgression, mostly introduced alleles were likely to have been
selected against in humans \citep{harris2016genetic, juric2016strength}. Since
Neanderthal and Denisovan population sizes were relatively small for hundreds
of thousands of years, theory predicts they would have accumulated fixed
deleterious variation at an increased rate. Introgressed haplotypes loaded with
more deleterious mutations would then have quickly been selected against after
admixture. Mapping the distribution of Neanderthal-introgressed haplotypes in
humans shows a reduction of Neanderthal-related ancestry in coding and
regulatory regions \citep{petr2019limits, telis2020selection,
yermakovich2023long}. These ``deserts'' of Neanderthal ancestry support the
hypothesis that introgressed functional alleles were selected against
\citep{sankararaman2014genomic, sankararaman2016combined}.

There is growing genetic evidence that early \emph{H. sapiens} reciprocally
contributed to Neanderthal genomes \citep{kuhlwilm2016ancient,
hubisz2020mapping, harris2023diverse}. Such introgression would have occurred
tens or hundreds of thousands of years prior to Neanderthal introgression in
humans during the global dispersal of modern humans around 60 ka. This is
supported by ``near modern'' \emph{H. sapiens} outside of African around
120--100 ka or earlier \citep{schwarcz1988esr, grun2005u, beyer2021climatic},
potentially overlapping with Neanderthals and providing opportunities for early
contacts. While estimates of the genomic contribution of early \emph{H.
sapiens} to Neanderthals vary, around $6\%$ of later Neanderthal genomes may
have been contributed through introgression from humans to Neanderthals
\citep{harris2023diverse}. If \emph{H. sapiens}-related haplotypes carried
fewer deleterious alleles due to their larger long-term effective population
size, human-introgressed DNA would have been favored in Neanderthal genomes.
The replacement of Neanderthal mitochondrial and Y chromosomes by early human
haplotypes supports this model of post-admixture selection in the Neanderthal
lineage \citep{posth2017deeply, petr2020evolutionary}.

Models for selection against introgressed alleles are often based on load
arguments (in particular, differences in the rate of accumulation of
unconditionally deleterious variation in populations of different sizes) or
hybrid incompatibilities \citep{muller1942isolating}. These arguments, founded
in population genetics theory, rarely take into account selection operating on
phenotypic variation or the relationship between genetic and phenotypic
variation. Many phenotypic traits are thought to be under stabilizing selection
\citep{sanjak2018evidence, sella2019thinking}, including gene regulation
\citep{gilad2006natural, hodgins2015gene, price2022detecting}. Because some of
the strongest signals of selection against Neanderthal-introduced alleles are
in regulatory regions \citep{sankararaman2014genomic}, stabilizing selection on
quantitative traits is particularly relevant to the dynamics of functional
genetic variation after introgression.

Stabilizing selection acts to maintain the phenotypic distribution of a trait
near some optimum, which is achieved by reducing phenotypic variation
(Figure~\ref{fig:stab-sel}A). When the mean phenotype of the population is
close to the phenotypic optimum, classical models predict that trait-affecting
alleles are subject to underdominant selection, i.e., selection against the
minor allele at the locus \citep{robertson1956effect}. This has proven to be a
useful model for understanding genetic architectures of traits under
stabilizing selection in single-population settings
\citep[e.g.,][]{keightley1988quantitative, simons2018population,
hayward2022polygenic}.

\begin{figure}[tb!]
    \centering
    \includegraphics{../figures/stab_sel.pdf}
    \caption{
        \textbf{Additive genetic variance under stabilizing selection.}
        (A) Stabilizing selection acts to maintain phenotypic values of individuals
        in the population near some optimum. Throughout, we assume a Gaussian fitness
        function.
        (B) With low mutational variance ($V_M$), the expected additive genetic
        variance ($V_G$) is proportional to the population-scaled mutation rate.
        When $V_M$ is large so that mutational effects can be strong, $V_G$ is
        independent of the mutation rate. The stochastic house-of-cards model
        (Eq.~\ref{eq:SHC}) interpolates these regimes, assuming steady-state
        dynamics \citep{burger1989much}. Expected $V_G$ computed using an SFS
        \citep[using \moments,][]{jouganous2017inferring} matches
        simulations assuming no linkage between loci affecting the trait, which
        align closely the stochastic house-of-cards approximation. Here, the
        mutation rate $\mu=0.01$ (per haploid), $V_S=1$, and $N_e=10,000$.
        Mutation effects are normally distributed with mean zero and variance
        $V_M$.
    }
    \label{fig:stab-sel}
\end{figure}

In diverged populations, the genetic variation contributing to a trait under
stabilizing selection has a higher rate of turnover compared to neutrally
evolving loci \citep{yair2022population}. We might therefore expect a rapid
divergence of trait architecture and an accumulation of fixed differences in
the \emph{H. sapiens} and Neanderthal lineages at trait-contributing loci, even
when the mean phenotype in each population remains close to the same trait
optimum. When a derived allele at high frequency in one population is
introduced to another population in which is was previously absent, it will be
at low frequency (if the admixture proportion is low) and subsequently selected
against. Likewise, if the ancestral allele is reintroduced to a population
fixed for a derived allele, the ancestral allele will be at low frequency and
will be selected against. In either case we should expect selection against the
introgressed allele, whether ancestral or derived and regardless the historical
relative sizes of the populations involved. This prediction contrasts with the
population-genetics ``load'' model, in which haplotypes with fewer deleterious
variants (such as those from the population with larger historical size) are
favored after introgression in either direction.

In this article, we show that admixture between diverged populations results in
selection against the minor parental ancestry at trait-affecting loci. We
develop a numerical approach based on the site frequency spectrum to predict
genetic variation under complex demographic scenarios, and which we use to
partition predicted trait heritability by introgressed vs. non-introgressed
variation. Using simulations with linkage, we demonstrate that deserts of
introgressed ancestry form around trait-contributing loci. When gene flow
occurs bidirectionally, such deserts will tend to overlap in location in each
population. As we argue in the Discussion, stabilizing selection on shared
trait optima provides an explanation for the overlap of introgressed ancestry
deserts in human and Neanderthal genomes after reciprocal introgression,
recently observed in \citet{harris2023diverse}.

\section*{Model and Methods}

We consider a polygenic trait for which an individual's additive genetic value
is the sum over all effects of alleles in their genome. For individual $i$,
\(G_i = \sum_l g_{l, i} a_l\), where $a_l$ is the effect size of the derived
allele at locus $l$, and \(g_{l,i}\in\{0,1,2\}\) is their genotype at that
locus (i.e., the number of alleles they carry at that locus). With no linkage
between trait-affecting loci, the expected additive genetic variance is \(V_A =
\sum_l 2p_l(1-p_l)a_l^2\), where $p_l$ is the allele frequency at locus $l$. We
ignore dominance and epistatic effects, so that \(V_G = V_A\). We further
ingore environmental effects \citep{simons2018population}, so that the
phenotypic variance \(V_P=V_G\).

Stabilizing selection acts to reduce phenotypic variation around the optimum
value $O$ (typally set to zero), and we assume a Gaussian fitness function
(Figure~\ref{fig:stab-sel}A) so that relative fitness is given by \(f(G_i | O,
V_S) = \exp{(-(G_i - O)^2 / 2 V_S)}\). $\frac{1}{V_S}$ is interpreted as the
strength of selection on the trait, so larger $V_S$ implies weaker selection.
For a population with mean phenotype at (or very close to) the optimum, the
mean fitness of the population (assuming a roughly normal distribution of
phenotypic values in the population) is
\[\bar{w} \approx \int_{-\infty}^\infty
f(G | 0, V_S) \mathcal{N}(0, V_G) dG =
\left(\frac{V_S}{V_S+V_G}\right)^{1/2}.\]
Thus, as the genetic variance increases, mean fitness among individuals in the
population decreases.

\subsection*{Mutation rates, effect sizes and genetic variance}

If all alleles contribute equally to the trait
(with effect sizes \(\pm a\) occurring in equal proportion),
\citet{keightley1988quantitative} showed that the dynamics of $V_G$ can be
approximated with the recursion
\[V_{G,t+1} \approx V_{G,t}\left(1-a^2/2(V_S+V_{G,t})\right)
\left(1-1/2N_e\right) + 2 \mu a^2,\]
where $\mu$ is the per-haploid, per-generation rate of mutation. In the
large-population-size limit, this gives the well-known result for
steady-state additive genetic variance,
\[V_G \approx 4\mu V_S,\] provided \(V_G \ll V_S\).

Mutations are not expected to each have the same effect size $|a|$, but rather
follow some distribution. Here, when mutation effects are not constant, we
assume effect sizes are drawn from a normal distribution with mean 0 and given
variance $V_M$. When population sizes or effect sizes are small, drift
dominates the dynamics of $V_G$, and at steady state \citet{lande1976natural}
found \[V_G \approx 4 N_e \mu V_M.\] Interpolating between the selection- and
drift-dominated regimes,
\begin{align}\label{eq:SHC}
    V_G & \approx \frac{4 \mu V_S}{1 + \frac{V_S}{N_e V_M}}.
\end{align}
This, the ``stochastic
house-of-Cards'' (SHC) approximation (Figure~\ref{fig:stab-sel}B), was given by
\citet{burger1989much} and is discussed in detail in \citet[][Ch.
28]{walsh2018evolution}.

\subsection*{Approximating allelic dynamics via underdominance}

As initially shown by \citet{robertson1956effect} \citep[see
also,][]{keightley1988quantitative, simons2018population}, when the mean
phenotype of the population is close to the optimal phenotypic value,
stabilizing selection results in symmetric underdominant selection on alleles
contributing additively to the trait. Thus, for an allele at frequency $p$ with
selection coefficient $s$, the expected change in $p$ over one generation is
\begin{equation}
    \E[\Delta p] = -sp(1-p)(1-2p).
\end{equation}
The selection coefficent $s$ depends on the strength of selection on the trait
$V_S$ and the effect size $a$, as 
\begin{equation}
    s\approx \frac{a^2}{2(V_S+V_G)} \approx \frac{a^2}{2V_S}.
\end{equation}
Details are shown in Appendix~\ref{sec:underdominance}.

\subsubsection*{Computing expected genetic variance from the SFS}

We model allelic dynamics using the diffusion approximation, where the expected
change in mean allele frequency per generation is
\[\E[M_{\delta_p}] = E[\Delta p] \approx -s p(1-p)(1-2p),\]
as shown above, and the expected change in variance of the allele frequency per
generation is
\[\E[V_{\delta_p}] \approx \frac{1}{2N_e}p(1-p).\]
Because $s$ is always positive for any $a\not=0$, we see that selection pushes
allele frequencies to zero if $p<1/2$ and to one if $p>1/2$, resulting in
symmetric underdominance.

We extend the \moments-based solution for the sample site-frequency spectrum
(SFS) \citep{jouganous2017inferring} to include underdominance with given
selection coefficient $s$ (Appendix~\ref{sec:moments-underdominance}). The
contribution of alleles with effect size $a$ to the total genetic variance
$V_G$ is found by computing the expected pairwise diversity from the SFS (with
sample size $n$, denoted $\Phi_n$), as \[V_{G,a} \approx 2a^2\sum_{j=1}^{n-1}
\frac{j(n-j)}{n(n-1)} \Phi_n(j|a,\mu_a) da.\] Here, $\mu_a$ is the mutation
rate of alleles with effect size $a$. Then assuming a normal distribution of
effect sizes for new mutations, the total genetic variance is \[V_G \approx
\int_{-\infty}^\infty V_{G,a}\mathcal{N}(0,V_M).\]

If $V_G$ is non-negligible compared to $V_S$, ignoring $V_G$ and using
$s=a^2/2V_S$ leads to deflated estimates of additive genetic variance
(Figure~\ref{fig:supp-one-pop}). When mutation rates are large so that $V_G$ is
not small compared to $V_S$, using $s=a^2/2(V_S+V_G)$ provides estimates of
$V_G$ that closely match simulations assuming free recombination between loci
(Figures~\ref{fig:toy-admixture} and
\ref{fig:one-pop-0.01}--\ref{fig:one-pop-0.1}). Because $V_G$ can vary over
time, this means that $s$ is no longer constant and can change due to factors
such as non-constant demography that increase or reduce $V_G$.

\begin{figure}[tb!]
    \centering
    \includegraphics{../figures/one_pop.pdf}
    \caption{
        \textbf{Population size changes and additive genetic variance.}
        We used \moments to track expected genetic variance of a trait
        under stabilizing selection as the population chages size. (A) The
        population goes through a 10-fold size reduction, followed by a recovery.
        (B) All mutations had effect sizes $\pm0.01$, $0.04$, or $0.1$, with equal
        probabilitiy of being trait increasing or decreasing. Depending on the
        selection coefficient ($s\approx a^2/2V_S$) compared to $N_e$, $V_G$
        could increase or decrease after a sudden population size change.
        Predictions using \moments match simulations with free recombination
        between trait-affeting loci
        (Figures~\ref{fig:one-pop-0.01}--\ref{fig:one-pop-0.1}).
    }
    \label{fig:one-pop}
\end{figure}

\subsubsection*{Demographic history}

Our numerical solution for the SFS allows for non-constant population size
histories, population splits, continuous migration and admixture. Here, we
consider relatively simple scenarios involving population splits with
subsequent introgression events. We focus on parameter regimes relevant to
human-Neanderthal history. In a simple toy model (which is not meant to
perfectly match any particular known or inferred history, but instead
demonstrate the effect of reciprocal introgression following population
divergence) a population of size $N_e=10{,}000$ splits into two, one remaining
size $10{,}000$ and the other shrinking to size $1{,}000$. They remain isolated
for $2N_e$ generations (or 500 thousand years, assuming an average generation
time of 25 years) and then introgression occurs from one branch to the other,
contributing $5\%$ ancestry to the recipient population
(Figure~\ref{fig:toy-admixture}A,B).

The second model is meant to more closely resemble inferred human-Neanderthal
history, in which the ancestral population of size $N_e=10{,}000$ splits into the
human and Neanderthal branches. At 250ka, an early human-to-Neanderthal
introgression contributes $5\%$ ancestry to Neanderthals. The human branch
shrinks to size $1{,}000$ 60ka, followed by exponential growth to size $20,000$
at present time. Neanderthals contribute $2\%$ ancestry to this bottlenecked
and expanding human population at 50ka, after which they go extinct
(Figure~\ref{fig:human-neand-h2}A).

In each scenario, we track genetic diversity and phenotypic variance before and
after admixture to understand allele frequency and haplotype dynamics. The
trait optimum is kept at $0$ in all branches, so that no optimum shift occurs,
and the strength of selection $V_S=1$ remains constant.

\begin{figure}[tb!]
    \centering
    \includegraphics{../figures/reciprocal_admixture.pdf}
    \caption{
        \textbf{Genetic variance of a trait under stabilizing seletion with
        multi-population demography.}
        (A,B) Two simple demographic models, in which Deme 1 has size 10,000
        and Deme 2 has size 1,000. In each scenario, we allow $5\%$ admixture
        from one deme to the other after begin isolated for 20,000 ($2N_e$)
        generations.
        (C,D) We compared predicted (additive) $V_G$ from the site frequency
        spectra (using \moments) to simulations
        without linkage between trait-affecting loci. In Deme 2, $V_G$ decreases
        after divergence and population size reduction. After admixture in both
        directions, $V_G$ increases and then decays again toward levels at steady
        state. Predictions from \moments closely match observed $V_G$ in
        simulations.
        Here, mutations were drawn from a normal distribution of fitnesses
        with $V_M=0.0025$.
        Other parameters: $\mu=0.025$, $V_S=1$, and the optimal phenotype remained
        the same in each population.
        (E,F) Using \moments, we partitioned additive contributions to $V_G$ by
        mutations that were previously segregating in the focal population,
        introgressed variants (either derived or reintroduced ancestral alleles),
        and by new mutations since the time of admixture. In both demographic
        scenarios, introgressed variants contribute a substantial proportion to
        $V_G$, though it is primarily composed of introduced derived alleles when
        admixture is from the small to the large population, while primarily
        reintroduced ancestral alleles under the reverse direction of gene flow.
        In both cases, $V_G$ is increasingly due to new mutations and the genetic
        architecture of the trait turns over with time since admixture.
    }
    \label{fig:toy-admixture}
\end{figure}

\subsubsection*{Simulations with free recombination}

We compare our \moments-based predictions for $V_G$ in non-equilibrium settings
to simulations assuming linkage equilibrium as well as individual-based
simulations with recombination (next section). For both, we allow mutations
to arise at rate $mu$ per haploid genome copy per generation. These are drawn
from some distribution of mutation effects, which then contribute to individual
trait values.

To simulate allele frequency changes under free recombination (assuming linkage
equilibrium between all trait-affecting alleles at all times), for each focal
segregating allele we integrate over possible genetic backgrounds contributed
by all other segregating alleles. We make the assumption that many alleles
contribute to the trait, so the variance in genetic backgrounds is normally
distributed around the mean genetic value $E_{\tilde{G}}=\sum_l 2p_l a_l$ with
variance $V_{\tilde{G}}=\sum_l 2p_l(1-p_l)a_l^2$, where the sums omit the focal
locus. The expected change in frequency is then computed using the approach
outlined in Appendix~\ref{sec:underdominance} (without taking the first-order
Taylor series approximation for the exponentials). Allele counts in the next
generation are then binomially sampled with parameter $p+\E[\Delta p]$
independently for each allele.

\subsubsection*{Individual-based simulations with linkage}

To include the effects of linkage between multiple selected alleles or selected
and neutral alleles, we used \fwdpy \citep{thornton2019polygenic} to run
Wright-Fisher simulations under the same demographic models considered above.
In these simulations, we consider large (1 Morgan, or 100 Mb with a per-base
recombination rate of $10^-8$) chromosomes with a uniform recombination
landscape and specify per-haploid mutation rates. Mutations may be uniformly
distributed across the chromosome, as in Figures~\ref{fig:linkage} and
\ref{fig:supp-low-VM}--\ref{fig:supp-high-VM}, or they may fall within
functional regions, as in Figures~\ref{fig:deserts} and
\ref{fig:deserts-directional}. In these simulations, such regions are centered
2 Mb apart and are 100kb in size, so that there are 50 evenly spaced regions
across the chromosome. Mutation effect sizes are drawn either as constant
values $\pm a$, or from a normal distribution with given mutational variance
$V_M$.

In these simulations, we track the empirical phenotypic variance (since there
is no simulated environmental effect, this is equivalent to $V_G$) in each
population each generation. We measure the effects of linked selection on
(neutral) introgressed ancestry using \tskit to analyze genealogical
information \citep{ralph2020efficiently}. To measure the reduction in
introgressed ancestry around introgressed variants (as shown in
Figure~\ref{fig:linkage}), we find each locus with a fixed difference between
the parental populations at the time of admixture by preserving the generation
immediately preceding admixture. For such a locus, for each sample we determine
which (preserved) parental population its ancestry traces to at varying
distances from the selected locus. Ancestry proportions are then averaged over
each fixed difference. To measure proportions of introgressed ancestry or
probabilities of observed ancestry deserts (as in Figure~\ref{fig:deserts}), we
take the same approach to average ancestral source population proportions in 50
kb windows. Ancestry deserts are defined as any such window with no ancestry
inherited from the source population of introgression events.

\subsection*{Data and code availability}

All code to run analyses, create figures, and compile this manuscript are available
at \url{https://github.com/apragsdale/neanderthal_stabilizing_selection}.

\section*{Results}

\subsection*{Additive genetic variance after admixture}

We expect genetic variance to increase after introgression. The amount that
genetic variance increases depends on the allelic differences accumulated
between populations and the effects of those alleles. Assuming no linkage,
dominance or epistasis, \(V_G=\sum_l 2p_l(1-p_l)a_l^2\). After admixture, with
proportion $f$ contributed by the source population (labeled 0) into the focal
population (labeled 1), \(p_l = fp_{l,0} + (1-f)p_{l,1}\). Plugging into the
expression for $V_G$ and after some simple algebra
(Appendix~\ref{sec:VG-admixture}), we can express the expected genetic variance
directly after admixture as \[V_G = fV_{G,0} + (1-f) V_{G,1} + 2f(1-f)\sum_l
F_{2,l} a_l^2,\] where $F_2 = (p_0 - p_1)^2$ is the squared difference in
allele frequencies at a locus \citep{peter2016admixture}. This result is known
\citep[e.g.,][]{tufto2000quantitative}, showing that additive variance is equal
to that in the source populations weighted by their contributions, plus a term
that depends on the divergence at trait-affecting loci between the populations
weighted by the quadratic factor $2f(1-f)$.

$F_2$ at a given locus depends on the demographic history
relating the two populations and the effect size at the locus due to selection
on the trait. In the infinitesimal limit, involving many loci each of
vanishingly small effect, dynamics at a given locus will be approximately
neutral, so that $F_2$ depends only on the demography. In this case,
\begin{align} \label{eq:VG-admixture}
    V_G & \approx f V_{G,0} + (1-f) V_{G,1} + 2f(1-f)\mathbb{E}[F_2] \sum_l a_l^2 \\
    \nonumber
    & = f V_{G,0} + (1-f) V_{G,1} + 2f(1-f)\mathbb{E}[F_2] L V_M,
\end{align}
where $L$ is the number of trait affecting loci.

\subsubsection*{Predicted $V_G$ from the SFS}

For $a \not\approx 0$, the assumption of neutral evolution will lead to
overestimates of $F_2$ compared to underdominant selection. We can compute
expected $V_G$ both before and after admixture using the diffusion
approximation for the joint SFS with underdominance. Comparing to simulations
with free recombination between loci (Methods), we find that this provides an
excellent fit to average observed $V_G$ over time
(Figure~\ref{fig:toy-admixture}C,D). In general, admixture causes a sudden
increase in $V_G$ followed by a fairly rapid return to pre-admixture levels.

Modeling the dynamics of genetic variance using the SFS lets us examine
contributions to $V_G$ from different classes of mutations, such as those at
different frequencies or arising at different times, and how those
contributions change over time. Of interest is the contribution to $V_G$ from
alleles that were already segregating in the recipient population, those that
were introduced through introgression, and new mutations since the time of
admixture (Figure~\ref{fig:toy-admixture}E,F). In many scenarios, the initial
divergence of the two populations will have occurred long enough before the
admixture event, so that the mutations underlying $V_G$ are largely unique in
each population. After mixing, previously segregating and introgressed alleles
go to fixation or loss, and the variance of the trait is increasingly due to
new mutations.

Introgressed variation can initially make up a considerable portion of $V_G$,
with those alleles being either newly introduced derived alleles or
reintroduced ancestral alleles. The relative sizes of the two populations
impacts the number of each, as derived alleles will accumulate more readily in
a population with smaller effective size. Nonetheless, the increase in $V_G$
can be fairly similar in either direction of introgrssion, as the term
\(2f(1-f)\sum_l F_{2,l} a_l^2\) contributes in either case and can be much
larger than $f V_G$ from the source population.

\subsubsection*{Complex demography and partitioning heritability by origin of alleles}

\begin{figure}[t!]
    \centering
    \includegraphics{../figures/h2-per-SNP.pdf}
    \caption{
        \textbf{Allele contributions to heritability under human-Neanderthal
        reciprocal introgression.}
        (A) A simple model of divergence and admixture between humans and
        Neanderthals. Using \moments, we computed predicted $V_G$ over time,
        partitioned by variation that was introgressed vs. non-introgressed
        (Figure~\ref{fig:human-neand-VG}).
        (B-E) Predicted per-SNP contributions to genetic variance ($h^2$ per
        SNP) is plotted over the 50 thousand years following introgression.
        For non-introgressed variants, we also plot $h^2$ per SNP weighted by
        allele frequencies matching those of introgressed variants. These
        are shown for (B,C) Neanderthal-to-human introgression 50 kya,
        (D,E) human-to-Neanderthal introgression 250 kya, (B,D) $V_M=0.0025$,
        and (C,E) $V_M=0.0001$.
    }
    \label{fig:human-neand-h2}
\end{figure}

We used a historical model more closely resembling inferred human-Neanderthal
history (Figure~\ref{fig:human-neand-h2}A) to explore the effects of population
size changes and reciprocal admixture on the additive genetic architecture of
traits under stabilizing selection. As expected
(Figure~\ref{fig:toy-admixture}), population contractions decrease $V_G$ as the
increased rate of drift reduces allelic diversity at trait-affecting loci, and
introgression increases $V_G$ in both directions of gene flow.

Because the genetic architectures considered here are purely additive, we can
track mutations in an admixed populations by whether they were previously
segregating, fixed or lost in either parental populations or if they arose as
new mutations since the time of admixture. This allows us to partition $V_G$ by
contributions from these different sets of mutations
(Figure~\ref{fig:human-neand-VG}). At low introgression proportions, as modeled
here, the majority of $V_G$ is still contributed by previously segregating,
non-introgressed mutations. $V_G$ due to existing mutations decays
monotonically over time and is fairly rapidly replace by new mutations.

The average contribution of introgressed vs. non-introgressed SNPs to $V_G$
(i.e., $h^2$ per SNP) can similarly by tracked over time. For the
human-Neanderthal-like demographic model and genetic architectures considered
here, the contribution per SNP of introgressed variants is initally lower than
that of non-introgressed variants. These contributions change over time,
depending on mutational variance, as well as demography
(Figure~\ref{fig:human-neand-h2}). When weighting $h^2$-per-SNP of
non-introgressed SNPs by matching to allele frequencies of introgressed
variants, relative contributions depend sensitively on evolutionary parameters
and the time since admixture.

\subsection*{The effects of linkage}

In the preceding sections, we find that approximating the dynamics of
trait-affecting alleles using an underdominance model
\citep{robertson1956effect} provides an excellent approximation of $V_G$ in
complex demographic scenarios. However, this relies on an absence of linkage
between trait-affecting alleles. The inclusion of linkage can lead to
noticeable effects on expected $V_G$, so that Equation~\ref{eq:SHC} differs
from observed $V_G$ at steady state
\citep{burger1989much, burger1994distribution, walsh2018evolution}.

To investigate the effects of linkage, we used chromosome-scale
individual-based simulations \citep{thornton2019polygenic}. By varying the
mutation rate and the variance of effect sizes of new mutations, we include
scenarios ranging from low to high polygenicity and from weak to strong
selection on individual alleles. In this and the following sections, we will
highlight two effects of linkage. First, we observe deviations of $V_G$ from
expectations without linkage, which can be large for highly polygenic traits.
Second, selection on introgressed trait-affecting alleles results in a
reduction of introgressed ancestry in surrounding regions.

With linkage between two or more alleles affecting a trait under stabilizing
selection, linkage disequilibrium (LD) can develop between alleles. Classicaly,
stabilizing selection will cause mutations of opposite-signed effects to be in
coupling LD and those of same-signed effects in repulsion LD
\citep{bulmer1971effect}. This has the effect of reducing $V_G$. For lower
mutational variances ($2N_e V_M\approx1$, Figure~\ref{fig:supp-low-VM}), we
observe this phenomenon. With low mutational input, and thus low polygenicity,
$V_G$ in individual simulations closely matches expectations without linkage.
As we increase the mutation rate, $V_G$ is reduced relative to those
expectations.  However, when the mutational variance is much larger
($2N_eV_m\approx 50$), we see the reverse trend
(Figure~\ref{fig:supp-high-VM}). At low mutation rate, there is a close match
between observed $V_G$ and unlinked expectations, although simulated values are
slightly higher. As the mutation rate increases, $V_G$ increases to be
relatively much larger than expectations, rather than smaller.

The strength of the deviation of $V_G$ between models with and without linkage
depends on a number of factors. The total mutation rate affects not only the
polygenicity, with higher mutation rates causing higher number of segregating
alleles, but also the mean recombination rate between those alleles, as they
will be more or less densely distributed in the genome. The distribution of
effect sizes plays an important role, as seen by the opposite trends in $V_G$
relative to unlinked predictions (Figure~X vs. Figure~Y), which are most
apparent with high mutation rates, and thus high polygenicity. In such cases,
there appear to be complex dynamics involving the Bulmer Effect
\citep{bulmer1971effect} and interference between selected alleles
\citep{hill1966effect}.

\begin{figure}[t!]
    \centering
    \includegraphics{../figures/ancestry_reduction.pdf}
    \caption{
        \textbf{Reduced introgressed ancestry around alleles contributing to a
        trait under stabilizing selection.}
        Using a deterministic model
        (Equations~\ref{eq:system-p}--\ref{eq:system-D}), we model the
        frequencies of an introgressed trait-affecting allele and a linked
        neutral allele, initially absent from the recipient population so that
        their frequencies equal the introgression proportion ($f=0.05$).
        The neutral allele frequency is reduced at a rate that depends on
        the effect size of the selected allele and the probability of
        recombination between them. LD (as measured by $D=Cov(p,q)$) decays
        to zero over time.
        (C) Alleles with strong effects are expected to result in a larger
        depletion of introgressed ancestry around the selected locus.
        (D,E) Compared to individual-based simulations, deterministic model
        predicts the dip in introgressed ancestry around trait-affecting loci,
        when the mutation rate is low, so that trait-affecting are sparsely
        distributed. When mutation rates (and thus polygenicity) are high,
        selected alleles are close together, so that selective interference
        is more pronounced and local ancestry is affected by multiple selected
        alleles.
    }
    \label{fig:linkage}
\end{figure}

\subsubsection*{Introgressed ancestry is reduced around introduced
trait-affecting alleles}

Because introgressed trait-affecting alleles are selection against as the minor
allele, introgressed ancestry segments in the regions surrounding the selected
alleles will also be removed due to linkage. The expected reduction in such
introgressed ancestry will depend on the effect size of the trait-affecting
allele and the probability of recombination surrounding that locus. We first
consider a deterministic model of the dynamics of introgressed allele and
linked ancestry frequencies with variable recombination between them. This
simple model ignores the effects of drift and of interference between multiple
selected alleles.

We model a trait-affecting site with a fixed difference between the two
parental populations, in which the derived allele may be fixed in either
population. At the functional site, with admixture proportion $f$ from the
minor parental source, the initial frequency of the derived allele is either
$p_0=f$ or $p_0=1-f$. Over one generation, the expected allele frequency at the
selected locus is, to leading order in $s$,
\begin{align}\label{eq:system-p}
    p_{t+1} & = p_t - s p_t(1-p_t)(1-2p_t),
\end{align}
with
\(s=a^2/2V_S\). As discussed above, if \(p_0<1/2\), \(p_t\rightarrow0\), and if
\(p_0>1/2\), \(p_t\rightarrow1\) as \(t\rightarrow\infty\).

We consider a neutral locus separated from the selected locus by recombination
probability $r$. Initially, the expected frequency of linked introgressed
ancestry is \(q_0=f\), which changes over time due to linked selection on the
trait-affecting allele. Letting \(D=Cov(p,q)\) be the standard covariance
measure of LD between the two loci, $q$ is expected to change as
\begin{align}\label{eq:system-q}
    q_{t+1} & = q_t - s D_t(1-2p_t).
\end{align}
Initially, \(D_0=\pm f(1-f)\), with \(D\) being
positive if \(p_0=f\) and negative if \(p_0=1-f\). LD between the loci changes
deterministically over time, due to both selection and recombination, so that
\begin{align}\label{eq:system-D}
    D_{t+1} = D_t - r D_t - s D_t (1-2p_t)^2.
\end{align}
Together, this forms a
simple nonlinear system of equations for the deterministic change in allele
frequencies at the two loci and LD between them.

Using this model, we predict the changes in introgressed allele frequences and
LD after admixture for given effect size $a$ and recombination rate $r$
(Figure~\ref{fig:linkage}A,B). As expected, smaller effect sizes result in a
slower decay in introgressed ancestry frequency at both selected and linked
loci, and larger recombination rates more quickly decouple the linked ancestry
from the selected allele dynamics. Thus, the expected reduction in introgressed
ancestry is largest for larger effect sizes (Figure~\ref{fig:linkage}C),
and LD between the selected and linked neutral alleles is largest for neutral
sites closest to the selected allele (Figure~\ref{fig:supp-LD}).

We assessed the accuracy of the deterministic two-locus model using
individual-based forward-in-time simulations \citep{thornton2019polygenic}
under a simple demographic model (Figure~\ref{fig:toy-admixture}A), with
introgression fraction \(f=0.05\). We simulated a single chromosome of length 1
Morgan, with all mutations having effect sizes \(\pm a\) (\(a=0.02\) or
\(0.05\)), and we varied the total per-chromosome mutation rate (\(\mu=0.001\),
\(0.0025\) and \(0.01\)). For each fixed-difference mutation in the parental
populations, we determined the average introgressed ancestry surrounding such
loci (Methods).

The deterministic model (Equations~\ref{eq:system-p}--\ref{eq:system-D})
provides a very good approximation when mutation rates are small. As the
mutation rate increases in these simulations, polygenicity also increases so
that trait-affecting alleles are more densely distributed along the chromosome.
Deviations from the deterministic model are due to multiple selected alleles
affecting local introgressed ancestry. For the highest mutation rate shown
here, there are expected to be multiple other trait-affecting alleles within a
1 cM window around any focal SNP, which interact with each other and also
affect local ancestry proportions.

\subsection*{Introgressed ancestry deserts are shared under stabilizing
selection and reciprocal introgression}

\begin{figure}[t!]
    \centering
    \includegraphics{../figures/introgression_deserts.SD_0.02.pdf}
    \caption{
        \textbf{Stabilizing selection causes an enrichment of introgressed
        ancestry deserts in functional regions.} (A) In chromosome-scale
        simulations under a demography with reciprocal migration between humans
        and Neanderthals (Figure~\ref{fig:human-neand-h2}A), stabilizing
        selection causes a chromosome-wide reduction of introgressed ancestry
        (below the $5\%$ and $2\%$ introgression proportions). This depletion
        is most pronounced in ``functional'' regions that allowed for
        trait-affecting mutations. (B) Introgressed ancestry deserts are
        more likely to occur in such functional regions, as are
        \emph{shared} deserts when compared across samples from humans and
        Neanderthals.
    }
    \label{fig:deserts}
\end{figure}

As shown in the previous section, introgressed ancestry is reduced around loci
with trait-affecting alleles regardless of the parental population the derived
allele is present in. We should therefore expect to observe reductions in
introgressed ancestry at the same trait-affecting loci after gene flow in
either direction. If introgression occurs in both directions, regions of
reduced introgressed ancestry will coincide around such loci, and introgression
``deserts'' that appear due to selection against trait-affecting alleles will
tend to be shared.

To demonstrate this effect, we performed chromosome-scale simulations under a
model of human-Neanderthal reciprocal introgression
\citep[Figure~\ref{fig:human-neand-h2}A,][]{harris2023diverse}. Mutations
affecting a trait under stabilizing selection in each population could arise in
100 kb ``functional'' regions, spaced 1 Mb apart (Methods). Sampling
individuals from both the human and Neanderthal lineages, we observed the
average introgressed ancestry proportions were lowest within functional regions
and dipped in the surrounding regions (Figure~\ref{fig:deserts}A). This
corresponded to an increased proportion of ancestry deserts (defined as
    contiguous 50kb regions with no observed introgressed ancestry in the
sample) within and immediately surrounding the functional regions
(Figure~\ref{fig:deserts}B). Functional regions also displayed an enrichment of
\emph{shared} ancestry deserts, with the probability of observing ancestry
deserts (either within a population or shared) decaying to background levels as
the distance from the functional region increases.

\section*{Discussion}

Many phenotypic traits are under stabilizing selection \citep{hodgins2015gene,
sanjak2018evidence, sella2019thinking}. This has motivated using stabilizing
selection around a shared optimum as a null model for the dynamics of alleles
affecting polygenic traits \citep{yair2022population}. Stabilizing selection
has a distinct effect on trait-affecting alleles, characterized by symmetric
underdominance \citep{robertson1956effect, keightley1988quantitative}, that is,
selection against the minor allele. Some theoretical and simulation studies
consider the effect of population differentiation and migration on the genetic
architecture of a trait under stabilizing selection
\citep[e.g.,][]{tufto2000quantitative, yeaman2011genetic, yair2022population},
but most previous work has focused on single population scenarios, often
assuming steady state dynamics. Admixture and introgression events are common
in natural populations, so understanding their effects on the architecture of
complex traits is needed.

Here, we have shown that multi-population, non-equilibrium approaches for the
site frequency spectrum with underdominance can be used to model the additive
genetic variance of a trait under stabilizing selection. While this approach
ignores biologically relevant aspects such as linkage between sites and
pleiotropy, it can still provide important insights into the dynamics of trait
architectures. By modeling allelic contributions to the trait as additive, it
allows for the decomposition of contributions from alleles of different
origins, either by source population or mutation time. It may therefore be
useful for understanding the expected contributions of introgressed variants to
complex traits, for example from Neanderthal-introgressed mutations in humans
\citep{wei2023lingering}. In the limited scenarios explored here, we see that
the expected contributions of introgressed variants to heritability are
complicated even in the purely additive case, depending on the distribution of
effect sizes, demographic history, and the time since admixture
(Figures~\ref{fig:human-neand-h2}, \ref{fig:human-neand-VG}).

After admixture between diverged populations, the additive genetic variance of
a trait rapidly increases. The genetic variance after admixture depends on both
the existing genetic variances within the parental population and their
divergence at trait-affecting loci, measured by $F_2$
(Equation~\ref{eq:VG-admixture}). $V_G$ is then expected to decay fairly
quickly to background levels. During this process, introgressed and
pre-existing trait-affecting alleles are replaced by new mutations, turning
over the genetic architecture of the trait (Figure~\ref{fig:toy-admixture}).
Selection occurs against the minor alleles at trait-affecting loci, so that
introgressed alleles from the minor parental ancestry, whether derived or
reintroduced ancestral alleles, are selected against
(Figure~\ref{fig:linkage}).

Selection against introgressed alleles will also remove linked introgressed
ancestry in the surrounding regions. Thus, introgressed ancestry deserts are
more likely to form at and around loci contributing to selected traits
(Figure~\ref{fig:deserts}). This process is symmetric, so that deserts tend to
form in the same regions under reciprocal introgression. When comparing
distributions of introgressed haplotypes in two diverged populations with more
recent gene flow, we should expect to see an overlap in deserts of introgressed
ancestry at loci affecting traits under stabilizing selection.

The expected pattern of shared ancestry deserts under stabilizing selection
differs from models of both deleterious load and incompatibilities. Load-based
models predict that haplotypes that have accumulated more deterious mutations,
e.g., from a population with small long-term effective population size, will be
selected against under either direction of gene flow. Introgressed ancestry at
a given selected locus will decrease in frequency in one introgression scenario
and increase the other. This may explain the replacement of MT and Y chromosome
DNA in Neanderthals by human haplotypes after early human-to-Neanderthal
introgression \citep{posth2017deeply, petr2020evolutionary}, but does not broadly
match observations across the autosomal genome \citep{harris2023diverse}.

The classical model of Bateson-Dobzhansky-Muller incompatibilities (BDMIs)
\citep{bateson1909heredity, dobzhansky1936studies, muller1942isolating}
explains the accumulation of reproductive isolation through negative epistatic
interactions that are exposed in hybrids. \citet{muller1942isolating}
hypothesized that such BDMIs should most often form between distant or unlinked
loci, instead of within single or tightly linked loci. Because theory predicts
that hybrid incompatibilities are resolved via selection against the minor
parental ancestry \citep{moran2021genomic}, ancestry deserts should form in
different genomic regions as the different incompatibility alleles are selected
against, depending on the direction of introgression. While such a model
predicts selection against introgressed ancestry, similar to that of
stabilizing selection, an important distinction is that deserts due to
selection against incompatibility loci are not expected to overlap under
bidirectional gene flow. While there is little empirical data on the
distribution of BDMIs, studies point to interacting BDMI alleles being unlinked
\citep[e.g.,][]{presgraves2003fine} and an asymmetry in the alleles under
selection in different introgression scenarios \citep{maheshwari2011genetics,
moran2021genomic}.

Finally, turning to the distribution of introgressed ancestry deserts in humans
and Neanderthals, \citet{harris2023diverse} find that regions of depleted
Neanderthal ancestry in humans overlap more than would be expected by chance
with regions lacking human-introgressed alleles in the Altai Neanderthal.
Human-introgressed ancestry in the Neanderthal genome is also depleted in
functional regions, as observed in humans \citep{sankararaman2014genomic}.
\citet{harris2023diverse} propose that epistatic interactions between
introgressed alleles and the recipient backgrounds could drive these patterns,
which they interpret as evidence for the inital process of speciation between
humans and Neanderthals. However, the observation of shared ancestry deserts
does not match expectations under a classic model of BDMIs as described above.
Instead, at least some of the pattern may be due to stabilizing selection
acting on complex traits, such as gene regulation. Importantly, such
overlapping ancestry deserts are expected even when a trait is under
stabilizing selection for the same phenotypic optimal value. The selective
causes of ancestry deserts in both humans and Neanderthals remain unknown.
Stabilizing selection will prove a useful null model for testing scenarios of
epistasis, incompatibilities and adaptive introgression.

%\section*{Acknowledgements}
%\begin{itemize}
%    \item Lab members
%    \item Kevin Thornton
%    \item Bret Payseur
%\end{itemize}

\bibliographystyle{genetics}
\bibliography{manuscript}

\appendix

\section{Stabilizing selection results in underdominant selection on
trait-affecting alleles} \label{sec:underdominance}

This result is well known and has been derived many times before. We include it
here for completeness. We consider a polygenic trait under stabilizing
selection, and we assume the mean phenotype of the population is close to the
optimal value (which will be the case if the optimum has not shifted recently).
Because many segregating loci are assumed to contribute to the phenotypic
variance of the trait, the phenotypic distribution is well-approximated as
normally distributed around the optimum: \(f(G)=\frac{1}{\sqrt{2\pi V_G}}
e^{-G^2/2V_G}\). We assume a Gaussian fitness function, so that the fitness
of an individual with genotypic value $G$ is given by \(w(G)=e^{-G^2/2V_S}\).

For a derived allele (labeled 1) with frequency $p$, the expected change in
frequency over one generation due to selection in general is
\[\E[\Delta p]=p(1-p)\frac{w_{1\cdot}-w_{0\cdot}}{\bar{w}},\]
where \(\bar{w}\) is the mean fitness of the population, and marginal fitnesses
for the derived and ancestral alleles are
\[w_{1\cdot}=p w_{11} + (1-p)w_{01},\]
and
\[w_{0\cdot}=p w_{01} + (1-p)w_{00},\]
respectively. Here, $w_{11}$, $w_{01}$, and $w_{00}$ are the relative fitnesses
of individuals who are homozygous for the derived allele, heterozygous, or
homozygous for the ancestral allele, respectively.

We partition an individual's genetic values into contributions from their
genetic background and the focal locus: $G=\tilde{G}+xa$, where $x\in{0,1,2}$.
We then integrate over genetic backgrounds to find the expected fitnesses of
each genotype. As described in \citet[][SI section 2.1]{simons2018population},
because each individual locus contributes a small amount to $V_G$,
$V_{\tilde{G}}\approx V_G$ and the fitness function is well approximated as
\(f(\tilde{G})=\frac{1}{\sqrt{2\pi V_G}}e^{-\tilde{G}^2/2V_G}\). Then,
mean fitness
\[
    \bar{w} = \sqrt{\frac{V_S}{V_S+V_G}},
\]
as shown in the main text, and
\[ 
    w_{11} \approx f(\tilde{G}) w(\tilde{G} + 2a) d\tilde{G}
    = \bar{w}\exp{\left(\frac{-4a^2(1-p)^2}{2(V_S+V_G)}\right)},
\]
\[
    w_{01} \approx f(\tilde{G}) w(\tilde{G} + a) d\tilde{G}
    = \bar{w}\exp{\left(\frac{-a^2(1-2p)^2}{2(V_S+V_G)}\right)},
\]
and
\[
    w_{00} \approx f(\tilde{G}) w(\tilde{G}) d\tilde{G}
    = \bar{w}\exp{\left(\frac{-4a^2p^2}{2(V_S+V_G)}\right)}.
\]

Taking the first-order Taylor series expansion of the exponentials
($e^{-x}\approx1-x$, for small $x$) and combining terms,
\[
    \frac{w_{1\cdot} - w_{0\cdot}}{\bar{w}} \approx \frac{-a^2(1-2p)}{2(V_S+V_G)},
\]
demonstrating the result.

\section{Moment equations for over- and underdominance}
\label{sec:moments-underdominance}

For symmetric underdominance, the relative fitnesses of genotypes \(aa:Aa:AA\)
are \(1:1-s:1\). We consider any value of $s$ (positive or negative,
corresponding to either under- or overdominance); with stabilizing selection,
\(s=\frac{a^2}{2V_S}\). We extended \moments \citep{jouganous2017inferring} to
compute the sample site frequency spectrum \(\Phi_{\mathbf{n}}\) for one or
more (up to five) populations with sample sizes \(\mathbf{n}\). This provides a
good approximation for the distribution of trait-affecting allele frequencies
across multiple populations if the trait optimum is shared across populations
and each population's mean phenotype remains close to that optimum. This is
expected to be the case if there are no optimum shifts in any lineage.
Hardy-Weinberg equilibrium is assumed at all loci. Accounting for optimum
shifts in one or more lineages would require a combination of direct and
underdominant selection \citep[e.g.,][]{hayward2022polygenic}, which we leave
for future work.

We refer readers to \citet{jouganous2017inferring} for a detailed introduction
to the general moments-based framework for the dynamics of \(\Phi_\mathbf{n}\).
Here, we describe how underdominant selection changes \(\Phi\) over a single
generation. With symmetric underdominance, selection acts heterozygotes. This
can be formulated as some proportion of heterozygotes failing to reproduce in a
given generation, with those selected lineages replaced by copies drawn from
the full population.

In a single population, we consider $n$ tracked lineages of which $i$ of those
carry the derived allele (\(\Phi_n(i)\) is thus the count of loci with $i$
observed derived alleles in a haploid sample of size of $n$). $i$ can increase
or decrease due to selection ``events''. Here, we assume $s$ is small enough so
that at most a single selection event occurs among the $n$ lineages in any
given generation. This is a reasonable approximation as long as $s$ is not
extremely large \citep{jouganous2017inferring} -- for selection coefficients
induced by stabilizing selection, this is typically a safe assumption.

Two selective events can change $i$ to $i+1$ or $i-1$: (a) a tracked copy
carrying a derived allele is heterozygous (paired with an ancestral
allele-carrying copy), selected against, and then replaced by an ancestral
allele drawn from the rest of the population, so that \(i\rightarrow i-1\), or
(b) a tracked copy carrying the ancestral allele is paired with a derived
allele-carrying copy, selected against, and then replaced by a derived allele,
so that \(i\rightarrow i+1\). In both cases, we require drawing two additional
lineages to find the \(\Phi_n\) in the next generation, one for the diploid
pair and one for the replacement allele (i.e., \(\Phi_n^{t+1}(i)\) requires
\(\Phi_{n+2}^t\)). This results in an unclosed system of equations, and we use
a quadratic jackknife approximation to approximate \(\Phi_{n+2}\) from
\(\Phi_n\), as described in \citet{jouganous2017inferring}.

For case (a), \(i\rightarrow i-1\) (i.e., \(\Phi_n(i)\) is reduced) with
probability \[s\frac{i(n-i+2)(n-i+1)}{(n+2)(n+1)}\Phi_{n+2}(i),\] and
\(i+1\rightarrow i\) (\(\Phi_n(i)\) increases) with probability
\[s\frac{(i+1)(n-i+1)(n-i)}{(n+2)(n+1)}\Phi_{n+2}(i+1).\] For case (b),
\(i\rightarrow i+1\) with probability
\[s\frac{(n-i)(i+2)(i+1)}{(n+2)(n+1)}\Phi_{n+2}(i+2),\] and \(i-1 \rightarrow
i\) with probability \[s\frac{(n-i+1)(i+1)i}{(n+2)(n+1)}\Phi_{n+2}(i+1).\]
These can be combined (with negative rates for the reduction of \(\Phi_n(i)\))
to describe the change of \(\Phi_n(i)\) for all \(0\leq i \leq n\).
Figures~\ref{fig:underdominance-validation-small-s} and
\ref{fig:underdominance-validation-large-s} show that this approach,
implemented in \moments, is accurate compared to discrete Wright-Fisher
simulations.

\section{Additive genetic variance after admixture}\label{sec:VG-admixture}

Here, we derive the expected genetic variance after admixture between two
source populations (Equation~\ref{eq:VG-admixture}). Suppose two population
(labeled 0 and 1) diverged some time in the past and then admix in proportions
$f$ and $1-f$.

Assuming no linkage, dominance or epistasis (so \(V_G=V_A\)),
\[V_G = \sum_l 2p_l(1-p_l)a_l^2 = \sum_l \pi_l a_l^2,\]
where $\pi$ denotes expected pairwise diverisity.
At a given locus (dropping the $l$), after admixture the allele frequency is
\[p=f p_0 + (1-f) p_1,\]
so that
\begin{align*}
    2p(1-p) & = 2(f p_0 + (1-f) p_1)(1 - f p_0 - (1-f) p_1) \\
    & = f^2 2p_0(1-p0) + (1-f)^2 p_1(1-p1) + 2f(1-f) (p_0(1-p_1) + p_1(1-p_0)) \\
    & = f^2 \pi_{0,0} + (1-f)^2 \pi_{1,1} + 2f(1-f)\pi_{0,1}.
\end{align*}

Plugging in to the definition for $V_G$, we get after admixture
\[V_G = f^2 V_{G,0} + (1-f)^2 V_{G,1} + 2f(1-f)\sum_l \pi_{0,1,l}a_l^2.\]
We can write $\pi_{0,1}$ at a given locus in terms of $\pi_{0,0}$, $\pi_{1,1}$,
and $F_2(0,1)=(p_0-p_1)^2$ as \citep{peter2016admixture}
\[\pi_{0,1} = F_2(0,1) + \frac{1}{2}\pi_{0,0} + \frac{1}{2}\pi_{1,1}.\]
Then
\begin{align*}
    V_G & = f^2 V_{G,0} + (1-f)^2 V_{G,1} + 2f(1-f)\sum_l \left[F_{2,l}(0,1)
    + \frac{1}{2}\pi_{0,0} + \frac{1}{2}\pi_{1,1}\right] a_l^2 \\
    & = f^2 V_{G,0} + (1-f)^2 V_{G,1} + 2f(1-f)\left[\frac{1}{2}V_{G,0} 
    + \frac{1}{2}V_{G,1} + \sum_l F_{2,l}(0,1)a_l^2\right] \\
    & = \left(f^2 + f(1-f)\right) V_{G,0} + \left((1-f)^2 + f(1-f)\right)V_{G,1}
    + 2f(1-f) \sum_l F_{2,i}(0,1)a_l^2 \\
    & = f V_{G,0} + (1-f)V_{G,1} + 2f(1-f)\sum_i F_{2,l}(0,1)a_l^2.
\end{align*}

\end{document}